---
title: "Gulf Stream related data"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The [gulfstream R package](https://github.com/BigelowLab/gulfstream) serves [Gulf Stream](https://www.nesdis.noaa.gov/about/k-12-education/oceans-coasts/what-the-gulf-stream) related data from a single package. All data are actively harvested and maintained using the [gstream R package](https://github.com/BigelowLab/gstream), but the [gulfstream R package](https://github.com/BigelowLab/gulfstream) is used for convenient serving of the same data in a portable package.

## Requirements

- [R v4.2+](https://www.r-project.org/)
- [rlang](https://CRAN.R-project.org/package=rlang)
- [dplyr](https://CRAN.R-project.org/package=dplyr)
- [tidyr](https://CRAN.R-project.org/package=tidyr)
- [ggplot2](https://CRAN.R-project.org/package=ggplot2)
- [lubridate](https://CRAN.R-project.org/package=lubridate)
- [sf](https://CRAN.R-project.org/package=sf)
- [leaflet](https://CRAN.R-project.org/package=leaflet)
- [shiny](https://CRAN.R-project.org/package=shiny)

## Installation

```
remotes::install_github("BigelowLab/gulfstream")
```

# Data

## Gulf Stream Index (GSI)

The [Gulf Stream
Index](https://en.wikipedia.org/wiki/Latitude_of_the_Gulf_Stream_and_the_Gulf_Stream_north_wall_index)
provides a positional index. Data are provides via the
[ecodata](https://noaa-edab.github.io/ecodata/) R package.

```{r gsi}
suppressPackageStartupMessages({
  library(dplyr)
  library(gulfstream)
})

read_gsi() |>
  glimpse()
```


## Gulf Stream SST Gradient Index (GSGI)

[Parfitt, Kwon, and Andres,
2022](https://agupubs.onlinelibrary.wiley.com/doi/10.1029/2022GL100914)
proposed a Gulf Stream Gradient Index. Data is served for 2004-2019
[here](https://www2.whoi.edu/staff/ykwon/data/).

> Parfitt, R., Y.-O. Kwon, and M. Andres, 2022: A monthly index for the
> large-scale sea surface temperature gradient across the separated Gulf
> Stream. Geophys. Res. Lett., 49, e2022GL100914.
> <https://doi.org/10.1029/2022GL100914>.

```{r gsgi}
read_gsgi() |>
  glimpse()
```


## Data from [RAPID](https://rapid.ac.uk)

### [RAPID-AMOC](https://rapid.ac.uk/rapidmoc)

Data from the RAPID AMOC monitoring project is funded by the Natural
Environment Research Council and are freely available from
www.rapid.ac.uk/rapidmoc.

Reference for Version v2020.2 
>Moat B.I.; Frajka-Williams E., Smeed
D.A.; Rayner D.; Johns W.E.; Baringer M.O.; Volkov, D.; Collins, J.
(2022). Atlantic meridional overturning circulation observed by the
RAPID-MOCHA-WBTS (RAPID-Meridional Overturning Circulation and Heatflux
Array-Western Boundary Time Series) array at 26N from 2004 to 2020
(v2020.2), British Oceanographic Data Centre - Natural Environment
Research Council, UK. <doi:10.5285/e91b10af-6f0a-7fa7-e053-6c86abc05a09>

```{r amoc}
read_moc_transports()|>
  glimpse()
```

### [RAPID-MOCHA](https://mocha.earth.miami.edu/mocha/data/index.html)

RAPID-MOCHA provides a heat transport time series.

```{r mocha}
read_rapid_mocha() |>
  glimpse()
```

## Gulf Stream Wall locations from US Navy

The daily data is hosted by by [NOAA’s Ocean Prediction
Center](https://ocean.weather.gov/) In particular they post the US
Navy’s [daily Gulf Stream point
data](https://ocean.weather.gov/gulf_stream_latest.txt) for the north
and south walls.

```{r usn}
read_usn()
```